---
title: "slope models"
author: "Leslie Jones"
date: "May 21, 2018"
output: html_document
---

```{r message=FALSE}
library(readr)
library(tidyverse)
library(ggplot2)
library(dplyr)
library(lubridate)
library(lme4)
library(nlme)
library(DT)


```
#Intro
Per Curry's email on April 18th:
Models are trying to address the question of whether streams with higher maximum temperatures are more likely to exhibit occupancy that is distributed further (to the right) from the thermal peak.

#Slope of Temperature
We fit a linear model to temperature time-series within the spawning window for each site-year combination. Theoretically,  slope of temperature experienced by population p in year t, or the rate of change in temperature at each site/year. 

#Temperature Metrics
We calculated Maximum temperature metrics for each site/year observation.

MxDAT = maximum daily average temp

MA7d_DAT = maximum 7 day average temp

sw_begin = julian day 1st day of spawn window


# Model idea
Model the relationship with a population-specific intercept, common effect of MaxTemp across populations, while treating your year t as a random effect on your intercept(s).


```{r}
facetSL <- read.csv("Data/Mergeslopefacets.csv")
facetSL <- facetSL %>% 
  select(site, year, MxDAT, MA7d_DAT, sw_begin, Intercept, Slope, sw_end)
datatable (facetSL)


```

```{r}
Mx <- lme(Slope ~ MxDAT + site  , random = ~1|year, data=facetSL)
summary(Mx)

MA7d_DAT <- lme(Slope ~ MA7d_DAT + site , random = ~1|year, data=facetSL)
summary(MA7d_DAT)
```

